---
title: "scrna clustering"
author: "Mike Mariani"
date: "4/19/2020"
output: html_document
---

```{r setup, include=FALSE}
##knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, include = TRUE, results="hide", fig.width=12, fig.height=16)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, include = TRUE, results="hide")
```

## Methods to detemine number of PCs

There are a few common methods available to help 
determine  the number of principle components 
to  use in downstream scrna cluster analysis. 
As a reminder, biological rational for cluster 
determination should take precadence over 
computational heuristics in most cases. 

Some methods:
1. Biological insight
2. eLbow plot - simplest easy to use - look for sharp elbow or bend. 
3. Jackstraw plot - produces a significance score (p-value)
4. Suerat "DimPlot()" and "DimHeatmap()" functions with "reduction='pca'". 
5. silhouette plot

## Load Seurat object

First we import a Seurat object that has already undergone
QC, normalization, variable features identification, scaling
and intital principle component (PC) determination 
- 'Seurat::RunPCA(seurat.object)'  

```{r load_seurat}

library(Seurat)
integrated <- readRDS(file = "/slipstream/home/mmariani/projects/silhouette_plot_example/vzv_vs_mock_integrated.rds")

```

## Elbow plot

```{r elbow, echo=FALSE}

Seurat::ElbowPlot(integrated)

```

No sharp elbow or bend here but note that plot levels off
between ~5-7 PCs.

## Jackstraw Plot

Now let's try a more quantitative JackStraw plot.
Note that a p-value is calculated for each PC.

```{r jackstraw}

integrated <- Seurat::JackStraw(integrated, num.replicate = 100)
integrated <- Seurat::ScoreJackStraw(integrated, dims = 1:20)
Seurat::JackStrawPlot(integrated, dims = 1:20)

```

The Jackstraw plot shows us that quite a few PCs could be inlcluded,
~4-7 might be a good number because these group largely to the left
of the plot but still hard to say definitively.

## DimPlot and DimHeatmap

Lets also take advantage of the Seurat DimPlot
and DimHeatmap functionality (make sure to 
specify "reduction='pca'" in this case)

```{r dimplot}

Seurat::DimPlot(integrated, reduction = "pca")
Seurat::DimHeatmap(integrated, dims=1:length(unique(integrated$seurat_clusters)), reduction="pca")

```

Hard to say.

With Dimplot() maybe about 4 distinct clusters.

With DimHeatMap() also a little tricky, PC8 and PC9 
start to look more uniform, so maybe we choose ~7

## Silhouette plot

Now let's try a silhouette plot to try and
mitigate some ambiguity.

```{r silhouette}

##Good information at:
##https://towardsdatascience.com/clustering-metrics-better-than-the-elbow-method-6926e1f723a6
##https://scikit-learn.org/stable/auto_examples/cluster/plot_kmeans_silhouette_analysis.html

##Silhouette code adapted from the following link (Stuart et al. 2019) :
##https://github.com/satijalab/Integration2019/blob/master/analysis_code/integration/integration_metrics.R#L36

library(cluster)

count <- 1
max_pcs <- 30 ##choose max number of PCs to test
sil.list <- list()
for(i in seq(from=1, by=1, to=max_pcs))
{
  reduction <- "pca"
  dims <- 1:i
  dist.matrix <- dist(x = Embeddings(object = integrated[[reduction]])[, dims])
  clusters <- integrated$orig.ident
  sil <- silhouette(x = as.numeric(x = as.factor(x = clusters)), dist = dist.matrix)
  ##summary(sil)
  sil <- sil[, 3]
  sil.list[[count]] <- sil
  count <- count+1
}

mean.sil.plot <- ggplot() +
  geom_point(aes(x=seq(from=1,by=1,to=max_pcs),y=unlist(lapply(sil.list,mean)))) +
  theme_bw() +
  ylab("mean silhouette coefficient") +
  xlab("number PCs")
mean.sil.plot

```

We see there are two "peaks"
after the initial down-slope.
Both peaks are close in magnitude,
so combining this information with the
other tests we have run, along with
pertinent biological insight, we 
conclude to use 7 PCs in our subsequent 
seurat clustering analyses.

## UMAP Clustering

```{r final, echo=FALSE}

integrated <- FindNeighbors(integrated, reduction = "pca", dims = 1:7)
integrated <- FindClusters(integrated, resolution = 0.5)
integrated <- RunUMAP(integrated, reduction = "pca", dims = 1:7)
integrated.dim.plot <- DimPlot(integrated, reduction = "umap", label=TRUE)
integrated.dim.plot

```
